---
title: "FinalProject"
author: "Sammit Bal, Jason Zheng, Thaddeus"
date: "2025-04-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Frontmatter

```{r}
library(glmnet)
library(rpart)
library(rattle) # for the fancyRpartPlot()
library(tidyverse)
library(readxl)

CODMaps <- read.csv("./CODMaps.csv")
CODGames_p2_380 <- read.csv("./CODGames_p2_380.csv")
CODGames_p1_380 <- read.csv("./CODGames_p1_380.csv")
CODGameModes <- read.csv("./CODGameModes.csv")


```

## Task 1: Data Cleaning and Data Visualization – Complete without Generative AI (Thaddeus + Sammit-ish)

> *Relevant Information:* (Complete without using Generative AI) Prior to each online match, players in the game lobby are presented with two options for the battlefield of the upcoming game (`Map1` and `Map2`). The players have the option to vote and the resulting vote is recorded in the `MapVote` column. The winning map is listed in the `Choice` column. In the event of a tie vote, the map listed in `Map1` is chosen. (Games for which the player entered the lobby after the vote has taken place have no information in `Map1` and `Map2` but have the winning map presented in `Choice`.)

**Research Question:** Which maps are the most likely to win the map vote when they are an option?

*Notes:* To answer this question...

1.  Write a paragraph (or more) discussing how you plan to answer this question (Be sure to address the data quality issues mentioned below and discuss how you will do the calculations).
2.  Then, write code and answer the question. (If I must answer questions about your approach/decision making process by reading your code rather than your discussion, you will lose points.)
3.  As part of your solution, you should calculate the proportion/probability that a map wins the vote given that it was a candidate.

*Note:* To do this, you will have to calculate the number times that each map was listed as a candidate (Map1 or Map2) and earned more votes than the other candidate.

-   As part of this, you should consider whether a given map won the vote by getting more votes than the other option or if it was selected since it was `Map1` and the vote was a tie.
-   You should also include a visualization of the results.
-   There might be some data quality issues (such as misspelled map names and extra (trailing) blanks in some entries) to solve for this problem.
-   You can find the proper names/spellings in the CODMaps.csv file.
-   **Warning:** To full receive full credit, you must write code to solve these issues rather than editing the `.csv` files

## Task 2: Data Cleaning and Data Visualization – Complete using Generative AI (Sammit + Thaddeus-ish)

**Question:** Repeat Task 1 using a generative AI of your choice.

1.  To answer this question, mention the tool (including version number if appropriate) you have selected.
2.  Then, discuss the prompt(s) you have used and provide the solution produced by the generative AI.

> *Note:* While it is fine to paste the question into the generative AI as your first prompt, you should also use additional follow-up prompts if it is beneficial to do so. - Be sure to discuss all prompts used in your report.

3.  Then, implement the generative AI solution.
4.  Finally, and most importantly, you should compare your solution from Task 1 to the generative AI solution.

> *Note 2:*
>
> -   Discuss similarities/differences, strengths/weaknesses, etc., and provide an overall assessment of which solution is better.
> -   The discussion should consider the correctness of the answers and should be substantial.
> -   Demonstrate that you have given the comparison considerable thought by making at least 3 substantial points as part of your comparison. - Each point should take the form of a well-written paragraph.

## Task 3: Inference (Sammit)

## Task 4 (Sammit)

